---
title: "p8105_hw1_ym3197"
author: "Yichaun Ma"
date: "2026-09-24"
output: github_document
---
```{r}
library(tidyverse)
library(ggplot2)
```
##Problem 1
import dataset
```{r}
data("penguins", package = "palmerpenguins")
```

The penguins data frame captures biological measurements from a sample of penguins. It records categorical information such as species, island, and sex, alongside continuous physical traits including bill length, bill depth, flipper length, and body mass. The dataset spans r nrow(penguins) observations and r ncol(penguins) variables. A key summary statistic is the mean flipper length, which evaluates to r mean(penguins$flipper_length_mm, na.rm = TRUE) %>% round(1) mm.
Now we create a scatter
```{r}
# Create the scatterplot
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  theme_minimal()
```
save the plot
```{r}

ggsave("penguin_plot.jpg")
```

## Problem 2
```{r}
df <- 
  tibble(
    sample_normal = rnorm(10),
    logical_vector = sample_normal > 0,
    character_vector = character(10),
    factor_vector= factor(c("small", "small", "small", "medium", "medium", "medium", "medium", "large", "large", "large"))
)
```

Try to take the mean 
```{r}
mean(pull(df, sample_normal))
mean(pull(df, logical_vector))
mean(pull(df, character_vector))
mean(pull(df, factor_vector))
```

sample_normal and logical_vector can be averaged because they are numeric and logical (0/1) vectors, respectively. Character and factor vectors, however, cannot be averaged. A potential solution might be to convert (coerce) these variables to a different data type.
```{r convert, results = 'hide'}

as.numeric(pull(df, sample_normal))
as.numeric(pull(df, character_vector))
as.numeric(pull(df, logical_vector))


```
This enables us to compute the mean for factor_vector since factor levels are mapped to integers based on their alphabetical sequence. Conversely, this approach fails for character_vector, as character vectors cannot be directly coerced into numeric values.